\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}

\title{Low Cycle Fatigue: Physics and Equations for Review}
\author{David Fieser \and Hugh Shortt\\[2pt]
\small lcf-strain-life. Both authors contributed equally.}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This is a self-contained statement of the fatigue physics implemented in the
toolkit, written for review by a materials or fatigue specialist. It contains the
equations, the symbols and units, the conventions, and the source for each
relation. It contains no software detail. A short sign-off table is at the end.
\end{abstract}

\section{Conventions}

All analysis uses true stress and true strain. Stress and modulus are in
\si{\mega\pascal}. Strain is a dimensionless fraction. Life is expressed in
reversals \(2N_f\), where one cycle is two reversals. The fatigue strength
exponent \(b\) and the fatigue ductility exponent \(c\) are negative.

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
Symbol & Meaning & Unit \\
\midrule
\(\sigma,\ \varepsilon\) & true stress, true strain & \si{\mega\pascal}, - \\
\(\Delta\sigma/2,\ \Delta\varepsilon/2\) & stress, total strain amplitude & \si{\mega\pascal}, - \\
\(\Delta\varepsilon_e/2,\ \Delta\varepsilon_p/2\) & elastic, plastic strain amplitude & - \\
\(\sigma_m,\ \sigma_{max}\) & mean stress, maximum stress & \si{\mega\pascal} \\
\(R,\ R_\varepsilon\) & stress ratio, strain ratio & - \\
\(E\) & Young's modulus & \si{\mega\pascal} \\
\(\sigma'_f,\ b\) & fatigue strength coefficient, exponent & \si{\mega\pascal}, - \\
\(\varepsilon'_f,\ c\) & fatigue ductility coefficient, exponent & - \\
\(K',\ n'\) & cyclic strength coefficient, strain-hardening exponent & \si{\mega\pascal}, - \\
\(2N_f\) & reversals to failure & - \\
\(K_t,\ K_f,\ q\) & stress concentration, fatigue notch factor, notch sensitivity & - \\
\bottomrule
\end{tabular}
\end{table}

\section{True stress and strain}
Engineering values \(e\) and \(\sigma_\text{eng}=F/A_0\) convert to true values by
\begin{equation}
\varepsilon = \ln(1+e), \qquad \sigma = \sigma_\text{eng}(1+e).
\end{equation}
Valid up to necking. Reference: Dowling, Mechanical Behavior of Materials, 4th ed.

\section{Cyclic stress-strain, Ramberg-Osgood}
\begin{equation}
\varepsilon = \frac{\sigma}{E} + \left(\frac{\sigma}{K'}\right)^{1/n'},
\qquad
\Delta\varepsilon = \frac{\Delta\sigma}{E} + 2\left(\frac{\Delta\sigma}{2K'}\right)^{1/n'}.
\end{equation}
The second form is the doubled hysteresis branch. Reference: Ramberg and Osgood
1943, NACA TN 902. Dowling 4th ed., Eq. 14.12.

\section{Strain-life}
Elastic, Basquin:
\begin{equation}
\frac{\Delta\sigma}{2} = \sigma'_f (2N_f)^{b},
\qquad
\frac{\Delta\varepsilon_e}{2} = \frac{\sigma'_f}{E}(2N_f)^{b}.
\end{equation}
Plastic, Coffin-Manson:
\begin{equation}
\frac{\Delta\varepsilon_p}{2} = \varepsilon'_f (2N_f)^{c}.
\end{equation}
Total strain-life and the elastic-plastic transition life:
\begin{equation}
\frac{\Delta\varepsilon}{2} = \frac{\sigma'_f}{E}(2N_f)^{b} + \varepsilon'_f (2N_f)^{c},
\qquad
2N_t = \left(\frac{\varepsilon'_f E}{\sigma'_f}\right)^{1/(b-c)}.
\end{equation}
Compatibility, checked and reported but not forced:
\begin{equation}
n' = \frac{b}{c}, \qquad K' = \frac{\sigma'_f}{(\varepsilon'_f)^{b/c}}.
\end{equation}
The plastic line is fit over the low cycle regime, excluding near-runout points
whose plastic strain is at measurement noise level. References: Basquin 1910,
Coffin 1954, Manson 1953, Dowling 4th ed. Eq. 14.3 to 14.6.

\section{Mean-stress corrections}
The asymmetry of a constant amplitude cycle is the stress ratio
\(R = \sigma_{min}/\sigma_{max}\), or the strain ratio
\(R_\varepsilon = \varepsilon_{min}/\varepsilon_{max}\) under strain control.
Strain-controlled testing is the norm in the low cycle regime per ASTM E606
and ISO 12106, stress control is the norm in the high cycle regime.

\begin{table}[h]
\centering
\small
\begin{tabular}{llll}
\toprule
Ratio & Loading & Minimum load & Maximum load \\
\midrule
\(R = -1\) & fully reversed, the LCF baseline & compression & equal tension \\
\(R = 0\) & pulsating tension & zero & tension \\
\(0 < R < 1\) & tension-tension & tension & higher tension \\
\(R = \pm\infty\) & pulsating compression & compression & zero \\
\(1 < R < \infty\) & compression-compression & higher compression & compression \\
\bottomrule
\end{tabular}
\end{table}

The control mode decides what a nonzero mean does over the life of a test.
Under strain control the mean stress relaxes toward zero as plastic strain
accumulates, and Morrow suits balanced or compressive means while SWT suits
tensile means. Under stress control the specimen ratchets, accumulating
strain in the direction of the mean stress, and the Walker fitted exponent
captures the stress ratio sensitivity. The corrections in this section apply
to the stabilized cycle. Reference: Dowling 4th ed.\ ch.~9.

The cycle-dependent evolution is modeled separately. Mean stress relaxation
under strain control follows \(\sigma_m(N) = \sigma_{m,1} N^{b_r}\) with
\(b_r \le 0\). Ratcheting under stress control accumulates strain as
\(\varepsilon_r(N) = C N^{p}\), and its life interaction is a
ductility-exhaustion penalty on the plastic strain-life line,
\(\Delta\varepsilon_p/2 = (\varepsilon'_f - \varepsilon_r)(2N_f)^{c}\). These
forms were reconstructed from collaborator notes and match the standard
published forms, pending the collaborator's confirmation. References: Morrow
and Sinclair 1958 (ASTM STP 237), Jhansale and Topper 1973 (ASTM STP 519),
Xia, Kujawski and Ellyin 1996 (Int.\ J.\ Fatigue 18:335), Kapoor 1994
(Fatigue Fract.\ Eng.\ Mater.\ Struct.\ 17:201).

Morrow, elastic term shifted by the mean stress:
\begin{equation}
\frac{\Delta\varepsilon}{2} = \frac{\sigma'_f-\sigma_m}{E}(2N_f)^{b} + \varepsilon'_f (2N_f)^{c}.
\end{equation}
Modified Morrow, both terms shifted:
\begin{equation}
\frac{\Delta\varepsilon}{2} = \frac{\sigma'_f-\sigma_m}{E}(2N_f)^{b}
+ \varepsilon'_f\left(\frac{\sigma'_f-\sigma_m}{\sigma'_f}\right)^{c/b}(2N_f)^{c}.
\end{equation}
Smith-Watson-Topper:
\begin{equation}
\sigma_{max}\,\frac{\Delta\varepsilon}{2} = \frac{(\sigma'_f)^2}{E}(2N_f)^{2b}
+ \sigma'_f \varepsilon'_f (2N_f)^{b+c},
\qquad
\sigma_{ar} = \sqrt{\sigma_{max}\,\sigma_a}.
\end{equation}
Walker, with the equivalent fully-reversed amplitude and the steel estimate of
the exponent:
\begin{equation}
\sigma_{ar} = \sigma_{max}^{\,1-\gamma}\,\sigma_a^{\,\gamma},
\qquad
\gamma_{steel} = 0.8818 - 2.00\times10^{-4}\,\sigma_u.
\end{equation}
Morrow equivalent fully-reversed amplitude:
\begin{equation}
\sigma_{ar} = \frac{\sigma_a}{1 - \sigma_m/\sigma'_f}.
\end{equation}
With \(\gamma=0.5\) Walker reduces to SWT. References: Morrow 1968. Smith, Watson,
Topper 1970, J. Materials 5(4):767. Walker 1970. Dowling, Calhoun, Arcari 2009,
Fatigue Fract. Eng. Mater. Struct. (steel \(\gamma\)). Dowling 4th ed. Eq. 9.18 to 9.21.

\section{Hysteresis energy and cyclic response}
The plastic strain energy density per cycle is the closed loop area
\begin{equation}
W = \oint \sigma\,d\varepsilon.
\end{equation}
The tension-compression asymmetry of a cycle is
\(R_{TC} = \lvert\sigma_{max}\rvert / \lvert\sigma_{min}\rvert\). Peak and valley
stress versus cycle give the cyclic hardening and softening response.

\section{Variable amplitude, cycle counting}
Irregular histories are reduced to closed cycles by the rainflow method, which
pairs reversals into hysteresis loops while preserving their order in time.
Reference: ASTM E1049-85(2017), Downing and Socie 1982, Matsuishi and Endo 1968.

Level-crossing counting records positive-slope crossings at and above a
reference level and negative-slope crossings below it. Peak counting records
peaks at and above the reference and valleys below it. Both follow ASTM
E1049-85(2017), sections 5.2 and 5.3.

The racetrack (gate) filter condenses a history before counting by removing
swings smaller than a gate while keeping the order of the large reversals.
Reference: Fuchs, Nelson, Burke, and Toomay, SAE paper 730565, 1973.

\section{Variable-amplitude local strain simulation}
A repeating strain history block is rotated to its global maximum and its
stress response simulated: the initial loading follows the cyclic
Ramberg-Osgood curve, every subsequent branch follows the doubled hysteresis
form given above from its reversal origin, and material memory follows the
rainflow closure rule, so a closed interior loop leaves the outer branch
exactly where it would have been without the interruption. Each closed loop
carries its strain amplitude and its simulated peak and mean stress, its
life comes from the SWT, Morrow, or uncorrected strain-life relation above,
and the damage is Miner-summed to blocks to failure. Sequence effects enter
through the loop means: a small cycle riding on a large branch carries the
mean stress of its position.

A load-input mode covers notched members: given a nominal stress history
and \(K_t\), the initial loading follows Neuber's rule on the cyclic curve
and every branch follows the modified Neuber rule on the doubled curve, so
the loops carry the local notch-root strain and stress.

Assumptions, stated plainly: stabilized cyclic properties throughout, no
cycle-dependent mean stress relaxation, no ratcheting. The simulation is
consistent with the constant-amplitude solvers and with rainflow counting by
construction and by test. Validation, strain input: against the published
SAE smooth-specimen dataset of Conle (MSc thesis, University of Waterloo,
1974, data distributed by the SAE FD and E committee), predictions fall
within a factor of two of experiment for the transmission and bracket
histories and about a factor of three, non-conservative, for the suspension
history, consistent with the documented scatter of linear-damage
local-strain predictions on this program. Validation, load input: against
the SAE keyhole benchmark (AE-6, Wetzel ed., 1977, inputs and results
preserved on the Internet Archive copy of the eFatigue benchmark page), the
constant-amplitude RQC-100 case predicts within 4 percent of the
benchmark's own strain-life calculation and conservatively against the
crack-based experimental life, and the Man-Ten suspension
variable-amplitude case predicts within a factor of two of the three
experimental lives. References: Masing 1926 (Proc.\ 2nd Int.\ Congress for
Applied Mechanics, Zurich), Dowling 4th ed.\ ch.~14, ASTM E1049-85(2017),
Neuber 1961, Conle 1974 and the SAE FD and E archive via fde.uwaterloo.ca.

\section{Cumulative damage}
Palmgren-Miner linear damage, failure at a critical sum, the default being one:
\begin{equation}
D = \sum_i \frac{n_i}{N_{f,i}}.
\end{equation}
Double Linear Damage Rule, Manson-Halford, with the Phase I life fraction
referenced to the longest life in the spectrum:
\begin{equation}
f_I = 0.35\left(\frac{N_f}{N_{long}}\right)^{0.25},
\qquad N_I = N_f f_I, \quad N_{II} = N_f(1-f_I).
\end{equation}
Phase I accumulates to one, then Phase II accumulates to one. Corten-Dolan, where
\(\sigma_1\) is the maximum stress in the block and \(\alpha_i\) the cycle fractions:
\begin{equation}
N = \frac{N_{f,1}}{\sum_i \alpha_i (\sigma_i/\sigma_1)^{d}}.
\end{equation}
References: Palmgren 1924, Miner 1945, Manson and Halford 1981 (Int. J. Fracture
17:169), Corten and Dolan 1956.

For stress-based collectives the allowable life comes from a one-slope Woehler
line with a knee at \((S_D, N_D)\):
\begin{equation}
N = N_D \left(\frac{S_a}{S_D}\right)^{-k} \quad (S_a \ge S_D),
\end{equation}
and below the knee one of three treatments: infinite life (Miner original),
the same slope \(k\) continued (Miner elementary), or the flatter fictitious
slope \(2k-1\) (Haibach):
\begin{equation}
N = N_D \left(\frac{S_a}{S_D}\right)^{-(2k-1)} \quad (S_a < S_D).
\end{equation}
References: Miner 1945, Haibach 1970, described in Haibach,
Betriebsfestigkeit, 3rd ed., Springer, 2006.

\section{Notch local-strain approach}
Neuber, combined with the cyclic curve, and its range form:
\begin{equation}
\frac{(K_t S)^2}{E} = \sigma\,\varepsilon,
\qquad
\frac{(K_t \Delta S)^2}{E} = \Delta\sigma\,\Delta\varepsilon.
\end{equation}
Glinka equivalent strain energy density:
\begin{equation}
\frac{(K_t S)^2}{2E} = \frac{\sigma^2}{2E} + \frac{\sigma}{n'+1}\left(\frac{\sigma}{K'}\right)^{1/n'}.
\end{equation}
Fatigue notch factor and notch sensitivity:
\begin{equation}
K_f^{\text{Peterson}} = 1 + \frac{K_t-1}{1+a/r},
\quad
K_f^{\text{Neuber}} = 1 + \frac{K_t-1}{1+\sqrt{\beta/r}},
\quad
q = \frac{K_f-1}{K_t-1}.
\end{equation}
Neuber tends to overestimate and Glinka to underestimate the local strain. The
measured value usually lies between them. References: Neuber 1961, Molski and
Glinka 1981, Peterson 1974.

\section{Statistics and design curves}
Life is the dependent variable in the linearized regression,
\(\log_{10} N = A + B \log_{10}(\Delta\varepsilon/2)\). Confidence and prediction
intervals use the residual standard error and the Student \(t\) quantile. A
reliability and confidence design curve is the mean reduced by \(k\,s\), where
\(k\) is the one-sided Owen tolerance factor. Right-censored runouts are handled
by maximum likelihood rather than deletion. References: ASTM E739, withdrawn 2024
and used as the de facto reference, Owen 1963, Williams, Lee and Rilly 2003 (Int.
J. Fatigue 25:427).

The fatigue limit from a staircase (up-and-down) test is estimated by the
Dixon-Mood method. With \(n_i\) the counts of the less frequent event on the
level grid, \(A=\sum i\,n_i\), \(B=\sum i^2 n_i\), \(N=\sum n_i\), step
\(d\), and \(X_0\) the lowest level where that event occurred,
\begin{equation}
\hat\mu = X_0 + d\left(\frac{A}{N} \mp \frac{1}{2}\right),
\qquad
\hat\sigma = 1.62\,d\left(\frac{NB-A^2}{N^2} + 0.029\right)
\quad\text{for}\quad \frac{NB-A^2}{N^2} \ge 0.3,
\end{equation}
with the minus sign when the analysis uses failures and the plus sign for
survivals. Below the 0.3 variability bound the standard deviation is the
approximate fallback \(\hat\sigma = 0.53\,d\) and is flagged. A-basis and
B-basis values are the one-sided lower tolerance bounds \(\bar x - k\,s\)
with the Owen factor at 99 percent reliability, 95 percent confidence and
90/95 respectively. When the data contain replicate amplitude levels the
linear fit is checked by the lack-of-fit F test,
\begin{equation}
F = \frac{SS_{LOF}/(m-2)}{SS_{PE}/(n-m)},
\end{equation}
where \(m\) is the number of distinct levels. References: Dixon and Mood
1948 (J.\ Amer.\ Statist.\ Assoc.\ 43:109), ISO 12107:2012, Owen 1963
(Sandia SCR-607), ASTM E739. Validated against the S34MnV staircase example
of Ekaputra et al.\ 2020 (Open Engineering 10:394).

The random fatigue limit model treats each specimen's fatigue limit
\(\gamma\) as unit-to-unit random, \(V = \log\gamma \sim
N(\mu_\gamma,\sigma_\gamma)\), with log life conditionally normal,
\begin{equation}
W = \log N \mid V \sim N\big(\beta_0 + \beta_1 \log(s-\gamma),\ \sigma\big),
\qquad s > \gamma,
\end{equation}
so the S-N curve flattens naturally near the limit and runouts enter the
likelihood as censored observations, including the probability that the
limit sits above the test stress. The five parameters are fit by maximum
likelihood with the marginal integral over \(V\) evaluated by quadrature.
Validation status, stated plainly: the implementation is verified against
brute-force integration and by recovering known parameters from data
simulated at the published laminate-panel design, it is not yet benchmarked
against the published fit itself because those raw datasets are not openly
published. Reference: Pascual and Meeker 1999 (Technometrics 41:277),
Meeker et al.\ 2026 (Statistical Science 41:1).

Outlier screening operates on the residuals of the log-life regression.
Runouts are censored observations, not outliers, and are excluded from the
screen. A single suspect point uses the two-sided Grubbs test, several suspect
points use the generalized extreme studentized deviate test, whose
approximation is intended for \(n \ge 15\). Influence is reported through
leverage, externally studentized residuals against a Bonferroni-corrected
\(t\) critical value, and Cook's distance against the \(4/n\) screening
threshold. Flagged points are reported, never deleted automatically.
References: Grubbs 1969 (Technometrics 11:1), Rosner 1983 (Technometrics
25:165), Cook 1977 (Technometrics 19:15), NIST/SEMATECH e-Handbook of
Statistical Methods, sections 1.3.5.17.1 and 1.3.5.17.3.

\section{Estimation of strain-life constants}
When no strain-controlled test data exists, the four constants are estimated
from monotonic properties. \(S_u\) is the ultimate tensile strength in
\si{\mega\pascal}, \(HB\) the Brinell hardness, \(RA\) the reduction in area
as a fraction, and \(\tilde\varepsilon_f = \ln[1/(1-RA)]\) the true fracture
ductility. Estimates are screening values, not substitutes for test data.

Medians method, steels (recommended default) and aluminum alloys:
\begin{equation}
\begin{aligned}
\text{steel:} &\quad \sigma'_f = 1.5\,S_u, \quad b=-0.09, \quad \varepsilon'_f = 0.45, \quad c=-0.59,\\
\text{aluminum:} &\quad \sigma'_f = 1.9\,S_u, \quad b=-0.11, \quad \varepsilon'_f = 0.28, \quad c=-0.66.
\end{aligned}
\end{equation}
Uniform Material Law, steels, with the ductility correction
\(\psi = 1\) for \(S_u/E \le 0.003\), else \(\psi = 1.375 - 125\,S_u/E\):
\begin{equation}
\sigma'_f = 1.50\,S_u, \quad b=-0.087, \quad \varepsilon'_f = 0.59\,\psi,
\quad c=-0.58, \quad K' = 1.65\,S_u, \quad n' = 0.15,
\end{equation}
and for aluminum and titanium alloys
\(\sigma'_f = 1.67\,S_u\), \(b=-0.095\), \(\varepsilon'_f = 0.35\),
\(c=-0.69\). The law loses validity as \(S_u\) approaches \SI{2.2}{\giga\pascal},
where \(\psi\) reaches zero.

Universal slopes, any metal:
\begin{equation}
\sigma'_f = 1.9\,S_u, \quad b=-0.12, \quad
\varepsilon'_f = 0.76\,\tilde\varepsilon_f^{\,0.6}, \quad c=-0.6.
\end{equation}
Modified universal slopes, steels:
\begin{equation}
\sigma'_f = 0.623\,E\left(\frac{S_u}{E}\right)^{0.832}, \quad b=-0.09, \quad
\varepsilon'_f = 0.0196\left(\frac{S_u}{E}\right)^{-0.53}\tilde\varepsilon_f^{\,0.155},
\quad c=-0.56.
\end{equation}
Hardness method, steels with roughly 150 to 700 HB:
\begin{equation}
\sigma'_f = 4.25\,HB + 225, \quad b=-0.09, \quad
\varepsilon'_f = \frac{0.32\,HB^2 - 487\,HB + 191000}{E}, \quad c=-0.56.
\end{equation}
References: Meggiolaro and Castro 2004 (Int. J. Fatigue 26:463, also the
comparative evaluation over 845 metals), Baeumel and Seeger 1990, Manson 1965
(Exp. Mech. 5:193), Muralidharan and Manson 1988 (J. Eng. Mater. Technol.
110:55), Roessle and Fatemi 2000 (Int. J. Fatigue 22:495).

\section{Elevated temperature}
Frequency-modified Coffin-Manson, coefficient form:
\begin{equation}
C_f = C_o\left(\frac{f}{f_{ref}}\right)^{k-1},
\qquad
\frac{\Delta\varepsilon_p}{2} = C_f (2N_f)^{c}.
\end{equation}
Linear time-fraction creep-fatigue damage, fatigue plus creep:
\begin{equation}
D = \sum_i \frac{n_i}{N_{f,i}} + \sum_j \frac{t_j}{t_{r,j}},
\end{equation}
checked against a bilinear creep-fatigue interaction envelope. References: Coffin
1971, Robinson 1952, ASTM E2714-13(2020).

\section{Multiaxial critical plane}
Critical-plane parameters:
\begin{equation}
P_{FS} = \frac{\Delta\gamma_{max}}{2}\left(1 + k\frac{\sigma_{n,max}}{\sigma_y}\right),
\quad
P_{BM} = \frac{\Delta\gamma_{max}}{2} + S\,\Delta\varepsilon_n,
\quad
P_{SWT} = \sigma_{n,max}\frac{\Delta\varepsilon_1}{2}.
\end{equation}
The plane search takes strain and stress tensor component histories sampled
over one cycle and scans plane normals over a hemisphere grid. Per plane,
the normal strain history is \(n\cdot\varepsilon\cdot n\) with amplitude
half its range, the shear amplitude is half the longest chord of the
in-plane shear-vector path, which stays meaningful for non-proportional
paths, and the maximum normal stress is the maximum of
\(n\cdot\sigma\cdot n\) over the cycle. Scope, stated plainly: amplitudes
come from the given cycle's path, per-plane rainflow counting of long
multiaxial histories is not implemented. Validated against the closed forms
for uniaxial loading with Poisson contraction (45 degree plane, engineering
shear amplitude \((1+\nu)\,\varepsilon_a\)) and pure torsion.
References: Fatemi and Socie 1988 (Fatigue Fract.\ Eng.\ Mater.\ Struct.\
11(3):149), Brown and Miller 1973, Smith, Watson, Topper 1970, Socie and
Marquis, Multiaxial Fatigue, SAE, 2000.

\section{Conventions and assumptions to confirm}
\begin{itemize}[nosep]
\item True stress-strain conversion assumed valid up to necking.
\item Plastic strain amplitude taken as \(\Delta\varepsilon_t/2 - \Delta\sigma/(2E)\).
\item Failure criterion is a percent load drop from the stabilized half-life peak load, default 30 percent.
\item Default mean-stress model for variable amplitude is SWT, notch default is Neuber, damage default is Miner.
\item The plastic strain-life line is fit over the low cycle regime only.
\end{itemize}

\section{Reviewer sign-off}
Reviewer name and date, then a verdict per section. Suggested verdicts: OK, OK
with note, or needs change.

\begin{table}[h]
\centering
\small
\begin{tabular}{p{6.5cm} p{2.5cm} p{5cm}}
\toprule
Section & Verdict & Note \\
\midrule
True stress and strain & & \\
Ramberg-Osgood & & \\
Strain-life & & \\
Mean-stress corrections & & \\
Energy and cyclic response & & \\
Cycle counting and filtering & & \\
Variable-amplitude simulation & & \\
Cumulative damage and knee variants & & \\
Notch local-strain & & \\
Statistics, design curves, outlier screen & & \\
Estimation of strain-life constants & & \\
Elevated temperature & & \\
Multiaxial critical plane & & \\
Conventions and assumptions & & \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
